---
title: "Markov Chains and Random Walks - HW9"
author: "Christian Thieme"
date: "10/27/2020"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1 - Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets 'A' dollars, he wins 'A' dollars with probability .4 and loses 'A' dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if:

* **A) He bets 1 dollar each time (timid strategy).**


```{r}
set.seed(123)

win <- 0
trials <- 100000


for (i in seq(1, trials)) {
  #we'll simulate the game 100,000 times
  
  smiths_money <- 1 #how much money smith has. He starts with $1
  
  
  while (smiths_money > 0 & smiths_money < 8) {   
    #while his money is greater than 0 and less than 8, keep playing the game
    
    outcome <-  sample(
        c(1, -1), #two outcomes, 1 or -1
        size = 1, #taking a single sample per turn
        replace = TRUE,
        prob = c(.4, .6) #.4 is the probability he wins
      ) 
    #using our probabilities, we'll take a single sample. 1 is a win, 
    #-1 is a loss
    
    smiths_money = smiths_money + outcome # current $ Smith has
    
    if (smiths_money == 8) {  #if he wins, we increment win
      win <- win + 1
    }
 
  }
  
}

win_probability <- win/trials

```

Smith's probability of winning with the 'timid strategy' is: `r win_probability`. 

<br>


* **B) He bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).**

```{r}
set.seed(123)

win <- 0
trials <- 100000
win_criteria <- 8


for (i in seq(1, trials)) {
  #we'll simulate the game 100,000 times
  
  smiths_money <- 1 #how much money smith has. He starts with $1
  
  
  while (smiths_money > 0 & smiths_money < win_criteria) {   
    #while his money is greater than 0 and less than 8, keep playing the game
    
    max_bet <- (win_criteria - smiths_money) #bets as much as possible but not
    #more than necessary
    
    if (smiths_money < max_bet) { 
        bet <- smiths_money #if money is less than the max bet, then he bets 
        #all his money
      } else  {
        bet <- max_bet #otherwise he only bets what he needs to
      }
      
    outcome <-  sample(
        c(bet, -bet), #two outcomes using the bet
        size = 1, #taking a single sample per turn
        replace = TRUE,
        prob = c(.4, .6) #.4 is the probability he wins
      ) 
    #using our probabilities, we'll take a single sample. 
    
    smiths_money = smiths_money + outcome # current $ Smith has
    
    } 
    
    if (smiths_money == 8) {  #if he wins, we increment win
      win <- win + 1
    }
    
  }
  

win_probability_2 <- win/trials
```

Smith's probability of winning with the 'bold strategy' is: `r win_probability_2`. 

<br>


* **C) Which strategy gives Smith the better chance of getting out of jail?**

Smith should use the 'bold strategy' because he's 3x more likely to get out of jail -- even though his chances are incredibly slim anyway. 
